---
title: "Adjusting income for inflation and cost of living"
author: "Andrew Ba Tran"
date: "October 14, 2016"
output: html_document
---

This walkthrough explains the methodology behind adjusting income in the Trend CT story [Would your job make you better or worse off in another state?](http://occupation.trendct.org/)

The data: 

* The annual median wage data comes from the Bureau of Labor Statistic's Employment and Wages from Occupational Employment Statistics (OES) [survey](http://stats.bls.gov/oes/). The data [goes back](http://www.bls.gov/oes/tables.htm) to 1988 but occupation codes changed around 1999, so started there to keep things as "apples to apples"" as possible.
* The inflation data is also from the Bureau of Labor Statistics' Consumer Price [index](http://www.bls.gov/cpi/).
* State-to-state cost of living adjustments was made by determining the relative value of $100 based on Bureau of Economic Analysis' Regional Price Parities for the [years available](http://www.bea.gov/newsreleases/regional/rpp/rpp_newsrelease.htm): 2008-2014.


## Loading the libraries

```{r setup, message=F, warning=F}
library(knitr)
library(lubridate)
library(dplyr)
library(blscrapeR)
```

## Loading the data

The BLS has their annual data in Excel spreadheets, which varies in structure year to year.

So it took some custom cleaning for each spreadsheet. 

If you want to see the specifics behind that, check out the `compiler.R` script. But it essentially creates two dataframes: `mega` and `titles`.

```{r data, message=F, warning=F}
mega <- readRDS("data/mega.rds")
titles <- read.csv("data/titles_codes_clean.csv")
kable(head(mega[,1:7]))
kable(head(titles))

## These are the column names of the combined BLS OES dataframe (also with some additional adjustments and rankings done in the compiler.R script)

colnames(mega)

## And how many years did we end up with?

range(mega$YEAR)

##  Let's take a look at some of the raw figures for median annual pay

raw_pay <- mega[c("STATE", "OCC_TITLE", "YEAR", "A_MEDIAN")]

## And look at a sampling of 2008 data

raw_tail <- subset(raw_pay, YEAR==2008)
raw_tail <- tail(raw_pay, 10)
kable(raw_tail)

## OK, let's adjust for inflation

## We'll use the blscrapeR package which is a wrapper for the BLS API.
### Here's the documentation https://cran.r-project.org/web/packages/blscrapeR/README.html


## This function pulls annual inflation rates compared to 2015
df <- inflation_adjust(2015)

kable(head(df))

## Adding a column specifically for year
df$year <- year(df$date)

## Narrowing down the dataframe
df <- df[c("adj_value", "pct_increase", "year")]

## Renaming the dataframe
colnames(df) <- c("adj_value", "pct_increase", "YEAR")

## Let's join it to our sample data frame

raw_tail <- left_join(raw_tail, df)

kable(raw_tail)

## What's the median pay adjusted for inflation? 

raw_tail$inflation_A_MEDIAN <- raw_tail$A_MEDIAN/raw_tail$adj_value

kable(raw_tail)

## So we know what the 1999 salaries were in 2015 dollars. 

## Next, let's adjust for power of the dollar in Wyoming compared to other states

value100s  <- read.csv("data/price_parities.csv", stringsAsFactors=F)

kable(head(value100s, 10))

## This is the regional price parity by state of $100

## Let's translate that to the how much $100 is worth in each state (Only for 2008)

value100s <- value100s[c("STATE", "y2008")]
kable(head(value100s))
```

```{r uh}
value100s$y2008_value <- 100+((100-value100s$y2008)/value100s$y2008*100)
value100s$y2008_percentchange <- (value100s$y2008_value-100)/100

kable(head(value100s, 10))
```
## Alright, lets join that to the inflation-adjusted dataframe from earlier
```{r}
raw_tail <- left_join(raw_tail, value100s)

kable(head(raw_tail, 10))

## And one more calculation to apply the price parity to Wyoming

raw_tail$adjusted_A_MEDIAN <- raw_tail$inflation_A_MEDIAN*raw_tail$y2008_percentchange+raw_tail$inflation_A_MEDIAN
raw_tail <- raw_tail[c("STATE", "OCC_TITLE", "YEAR", "A_MEDIAN", "inflation_A_MEDIAN", "adjusted_A_MEDIAN")]
kable(head(raw_tail, 10))

## So this is the process we took to determine the *actual* salary adjusting for year and state.

```

