---
title: "Story"
author: "Andrew Ba Tran"
date: "September 14, 2016"
output: html_document
---

```{r setup, include=FALSE}
library(readxl)
library(dplyr)
library(tidyr)
library(stringr)
wages2015 <- read_excel("data/state_M2015_dl.xlsx", sheet=1)
value100 <- read.csv("data/valueof100.csv", stringsAsFactors=F)
value100$per_change <- (value100$value100-100)/100
wages2015$OCC_TITLE <- str_to_lower(wages2015$OCC_TITLE)
wages2015$OCC_TITLE <- gsub(", general", "", wages2015$OCC_TITLE)

#wages2015$OCC_TITLE <- paste(toupper(substr(wages2015$OCC_TITLE, 1, 1)), substr(wages2015$OCC_TITLE, 2, nchar(wages2015$OCC_TITLE)), sep="")

wages2015 <- left_join(wages2015, value100)

wages2015$H_MEAN <- as.numeric(wages2015$H_MEAN)
wages2015$H_MEDIAN <- as.numeric(wages2015$H_MEDIAN)
wages2015$A_MEAN <- as.numeric(wages2015$A_MEAN)
wages2015$A_MEDIAN <- as.numeric(wages2015$A_MEDIAN)

wages2015$adjusted_H_MEAN <- wages2015$H_MEAN * wages2015$per_change + wages2015$H_MEAN
wages2015$adjusted_H_MEDIAN <-wages2015$H_MEDIAN * wages2015$per_change + wages2015$H_MEDIAN

wages2015$adjusted_A_MEAN <- wages2015$A_MEAN * wages2015$per_change + wages2015$A_MEAN
wages2015$adjusted_A_MEDIAN <-wages2015$A_MEDIAN * wages2015$per_change + wages2015$A_MEDIAN


wages2015 <- filter(wages2015, STATE!="Guam" & STATE!="Puerto Rico" & STATE!= "Virgin Islands")

wages2015a <- subset(wages2015, !is.na(A_MEAN))
wages2015b <- subset(wages2015, !is.na(A_MEDIAN))

wages2015a_raw <-  wages2015a %>%
  arrange(OCC_TITLE, -A_MEAN) %>%
  group_by(OCC_TITLE) %>%
  mutate(rank_A_MEAN=row_number()) %>%
  select(STATE, OCC_TITLE, rank_A_MEAN)


wages2015a_adjusted <-  wages2015a %>%
  arrange(OCC_TITLE, -adjusted_A_MEAN) %>%
  group_by(OCC_TITLE) %>%
  mutate(rank_adjusted_A_MEAN=row_number()) %>%
  select(STATE, OCC_TITLE, rank_adjusted_A_MEAN)

wages2015b_raw <-  wages2015b %>%
  arrange(OCC_TITLE, -A_MEDIAN) %>%
  group_by(OCC_TITLE) %>%
  mutate(rank_A_MEDIAN=row_number()) %>%
  select(STATE, OCC_TITLE, rank_A_MEDIAN)


wages2015b_adjusted <-  wages2015b %>%
  arrange(OCC_TITLE, -adjusted_A_MEDIAN) %>%
  group_by(OCC_TITLE) %>%
  mutate(rank_adjusted_A_MEDIAN=row_number()) %>%
  select(STATE, OCC_TITLE, rank_adjusted_A_MEDIAN)

wages2015 <- left_join(wages2015, wages2015a_raw)
wages2015 <- left_join(wages2015, wages2015a_adjusted)
wages2015 <- left_join(wages2015, wages2015b_raw)
wages2015 <- left_join(wages2015, wages2015b_adjusted)

#mega <- read.csv("mega_wages.csv", stringsAsFactors=F)

state_abbr<-"TX"
state_occ <-"Editors"
state_wages <- subset(wages2015, ST==state_abbr)
state_name <- unique(state_wages$STATE)
total_emp <-  subset(state_wages, OCC_TITLE=="All Occupations")
total_emp <- as.numeric(total_emp$TOT_EMP)
total_emp <- round(total_emp/1000000,1)

state_wages_median <- state_wages %>%
  arrange(A_MEDIAN)
state_wages_median <- subset(state_wages_median, !is.na(A_MEDIAN))

state_wages_mean <- state_wages %>%
  arrange(A_MEAN)
state_wages_mean <- subset(state_wages_mean, !is.na(A_MEAN))

state_wages_median_adjusted <- state_wages %>%
  arrange(adjusted_A_MEDIAN)
state_wages_median <- subset(state_wages_median, !is.na(adjusted_A_MEDIAN))

state_wages_mean_adjusted <- state_wages %>%
  arrange(adjusted_A_MEAN)
state_wages_mean <- subset(state_wages_mean, !is.na(adjusted_A_MEAN))

min_job_mean <- head(state_wages_mean, 5)
min_job_median <- head(state_wages_median, 5)

max_job_mean <- tail(state_wages_mean, 5)
max_job_median <- tail(state_wages_median, 5)

min_job_mean_adjusted <- head(state_wages_mean_adjusted, 5)
min_job_median_adjusted <- head(state_wages_median_adjusted, 5)

max_job_mean_adjusted <- tail(state_wages_mean_adjusted, 5)
max_job_median_adjusted <- tail(state_wages_median_adjusted, 5)

number1_median <- subset(state_wages, rank_A_MEDIAN==1)
number1_median_adjusted <- subset(state_wages, rank_adjusted_A_MEDIAN==1)

dol_value1 <- state_wages$value100[1]
state_number <- state_wages$AREA[1]
state_number2 <- as.numeric(state_number)
state_number2 <- ifelse(state_number2<26, state_number2+25, state_number2-25)
state_number2 <- ifelse(nchar(state_number2)==1, paste0("0", state_number2), as.character(state_number2))
state_number2 <- subset(wages2015, AREA==state_number2)
dol_value2_state <- state_number2$STATE[1]
dol_value2 <- state_number2$value100[1]


# Calculating disparity within occupations by state

occ_disparity <- wages2015 %>%
  group_by(OCC_CODE) %>%
  filter(!is.na(A_MEDIAN)) %>%
  arrange(A_MEDIAN) %>%
  filter(row_number()==1 | row_number()==n()) %>%
  mutate(TYPE=ifelse(row_number() %% 2 == 0, "Max", "Min")) %>%
  select(OCC_CODE, TYPE, STATE, A_MEDIAN) 

occ_disparity_max <- occ_disparity %>%
  filter(TYPE=="Max") %>%
  spread(TYPE, A_MEDIAN)

colnames(occ_disparity_max) <- c("OCC_CODE", "max_State", "max_Wage")


occ_disparity <- wages2015 %>%
  group_by(OCC_CODE) %>%
  filter(!is.na(A_MEDIAN)) %>%
  arrange(A_MEDIAN) %>%
  filter(row_number()==1 | row_number()==n()) %>%
  mutate(TYPE=ifelse(row_number() %% 2 == 0, "Max", "Min")) %>%
  select(OCC_CODE, TYPE, STATE, A_MEDIAN) 

occ_disparity_min <- occ_disparity %>%
  filter(TYPE=="Min") %>%
  spread(TYPE, A_MEDIAN)

colnames(occ_disparity_min) <- c("OCC_CODE", "min_State", "min_Wage")



occ_disparity_adj <- wages2015 %>%
  group_by(OCC_CODE) %>%
  filter(!is.na(adjusted_A_MEDIAN)) %>%
  arrange(adjusted_A_MEDIAN) %>%
  filter(row_number()==1 | row_number()==n()) %>%
  mutate(TYPE=ifelse(row_number() %% 2 == 0, "Max", "Min")) %>%
  select(OCC_CODE, TYPE, STATE, adjusted_A_MEDIAN) 

occ_disparity_max_adj <- occ_disparity_adj %>%
  filter(TYPE=="Max") %>%
  spread(TYPE, adjusted_A_MEDIAN)

colnames(occ_disparity_max_adj) <- c("OCC_CODE", "max_adj_State", "max_adj_Wage")


occ_disparity_adj <- wages2015 %>%
  group_by(OCC_CODE) %>%
  filter(!is.na(adjusted_A_MEDIAN)) %>%
  arrange(adjusted_A_MEDIAN) %>%
  filter(row_number()==1 | row_number()==n()) %>%
  mutate(TYPE=ifelse(row_number() %% 2 == 0, "Max", "Min")) %>%
  select(OCC_CODE, TYPE, STATE, adjusted_A_MEDIAN) 

occ_disparity_min_adj <- occ_disparity_adj %>%
  filter(TYPE=="Min") %>%
  spread(TYPE, adjusted_A_MEDIAN)

colnames(occ_disparity_min_adj) <- c("OCC_CODE", "min_adj_State", "min_adj_Wage")

occ_code_names <- wages2015[c("OCC_CODE", "OCC_TITLE")]
occ_code_names <- unique(occ_code_names)

occ_max_min <- left_join(occ_code_names, occ_disparity_max)
occ_max_min <- left_join(occ_max_min , occ_disparity_min)
occ_max_min <- left_join(occ_max_min , occ_disparity_max_adj)
occ_max_min <- left_join(occ_max_min , occ_disparity_min_adj)
occ_max_min$adj_diff <- occ_max_min$max_adj_Wage - occ_max_min$min_adj_Wage
#write.table(occ_max_min, "occ_max_min.csv", sep=",", na="")
#write.csv(occ_max_min, "occ_max_min.csv")


# Calculating disparity of occupation pay within states

state_disparity <- wages2015 %>%
  group_by(ST) %>%
  filter(!is.na(A_MEDIAN)) %>%
  arrange(A_MEDIAN) %>%
  filter(row_number()==1 | row_number()==n()) %>%
  mutate(TYPE=ifelse(row_number() %% 2 == 0, "Max", "Min")) %>%
  select(OCC_TITLE, TYPE, ST, A_MEDIAN) 

state_disparity_max <- state_disparity %>%
  filter(TYPE=="Max") %>%
  spread(TYPE, A_MEDIAN)

colnames(state_disparity_max) <- c("max_OCC_TITLE", "State", "max_Wage")

state_disparity_min <- state_disparity %>%
  filter(TYPE=="Min") %>%
  spread(TYPE, A_MEDIAN)

colnames(state_disparity_min) <- c("min_OCC_TITLE", "State", "min_Wage")



state_disparity_adj <- wages2015 %>%
  group_by(ST) %>%
  filter(!is.na(adjusted_A_MEDIAN)) %>%
  arrange(adjusted_A_MEDIAN) %>%
  filter(row_number()==1 | row_number()==n()) %>%
  mutate(TYPE=ifelse(row_number() %% 2 == 0, "Max", "Min")) %>%
  select(OCC_TITLE, TYPE, ST, adjusted_A_MEDIAN) 

state_disparity_max_adj <- state_disparity_adj %>%
  filter(TYPE=="Max") %>%
  spread(TYPE, adjusted_A_MEDIAN)

colnames(state_disparity_max_adj) <- c("max_adj_OCC_TITLE", "State", "max_adj_Wage")

state_disparity_min_adj <- state_disparity_adj %>%
  filter(TYPE=="Min") %>%
  spread(TYPE, adjusted_A_MEDIAN)

colnames(state_disparity_min_adj) <- c("min_adj_OCC_TITLE", "State", "min_adj_Wage")


state_max_min <- left_join(state_disparity_max , state_disparity_min)
state_max_min <- left_join(state_max_min , state_disparity_max_adj)
state_max_min <- left_join(state_max_min , state_disparity_min_adj)
state_max_min$max_adj_Wage <- as.numeric(state_max_min$max_adj_Wage)
state_max_min$min_adj_Wage <- as.numeric(state_max_min$min_adj_Wage)

state_max_min$adj_diff <- state_max_min$max_adj_Wage - state_max_min$min_adj_Wage

#write.table(occ_max_min, "occ_max_min.csv", sep=",", na="")
#write.csv(occ_max_min, "occ_max_min.csv")

# STATE PERCENTS EMPLOYEES
wages2015$TOT_EMP <- as.numeric(wages2015$TOT_EMP)

wages2015_no_totals <- wages2015 %>%
  filter(OCC_GROUP!="total")

wages2015_totals <- wages2015 %>%
  filter(OCC_GROUP=="total")

states <- unique(wages2015_no_totals$STATE)

for (i in 1:length(states)) {
  state_num <- states[i]
  states_subset <- subset(wages2015_no_totals, STATE==state_num)
  states_subset_total <- subset(wages2015_totals, STATE==state_num)
  states_subset_total_num <- as.numeric(states_subset_total$TOT_EMP[1])
  states_subset$per_employees <-round(states_subset$TOT_EMP/states_subset_total_num*100,2)
  states_subset_total$per_employees <- -1
  states_subset_all <- rbind(states_subset, states_subset_total)
  
  if (i == 1) {
    states_recombined <- states_subset_all
  } else {
    states_recombined <- rbind(states_recombined, states_subset_all)
  }
    
  
}


most_workers_detailed <- states_recombined %>%
  group_by(ST, OCC_GROUP) %>%
  filter(!is.na(TOT_EMP)) %>%
  arrange(-per_employees) %>%
  filter(row_number()==1) %>%
  filter(OCC_GROUP!="total" & OCC_GROUP=="detailed") %>%
  select(ST, OCC_TITLE, per_employees) 

colnames(most_workers_detailed) <- c("group", "STATE", "detailed_title", "detailed_percent")
most_workers_detailed$group <- NULL

most_workers_major <- states_recombined %>%
  group_by(ST, OCC_GROUP) %>%
  filter(!is.na(TOT_EMP)) %>%
  arrange(-per_employees) %>%
  filter(row_number()==1) %>%
  filter(OCC_GROUP!="total" & OCC_GROUP=="major") %>%
  select(ST, OCC_TITLE, per_employees) 

colnames(most_workers_major) <- c( "group", "STATE","major_title", "major_percent")
most_workers_major$group <- NULL

most_workers <- left_join(most_workers_major, most_workers_detailed)

most_workers_state <- subset(most_workers, STATE==state_abbr)
state_max_min_state <- subset(state_max_min, State==state_abbr)
```

# Focusing on `r state_name`

In `r state_name`, there are about `r total_emp` million full-time workers.

Most workers in the state are in major job category of `r most_workers_state$major_title` (about `r most_workers_state$major_percent` percent). More specifically, the most employed people in the state work as `r most_workers_state$detailed_title` (`r most_workers_state$detailed_percent` percent).

The annual median pay ranges from $`r formatC(round(min(state_wages_median$A_MEDIAN)), format="d", big.mark=',')` for `r min_job_median$OCC_TITLE[1]` at the lowest end of the spectrum to  $`r formatC(round(max(state_wages_median$A_MEDIAN)), format="d", big.mark=',')` for `r max_job_median$OCC_TITLE[5]` at the highest. That's a gap of $`r paste0(formatC(round(max(state_max_min_state$adj_diff)), format="d", big.mark=','),  ifelse(state_abbr=="MS", ", the largest in the country.", paste0(". Mississippi has the largest gap in the country of $193,000 between anesthesiologists and psychiatric aides.")))`

`r ifelse(substr(state_name, nchar(state_name), nchar(state_name))=="s", paste0(state_name, "'"), paste0(state_name, "'s"))` highest paid jobs include `r paste0(max_job_median$OCC_TITLE[2], ", ", max_job_median$OCC_TITLE[3], ", and ", max_job_median$OCC_TITLE[4])`.

About `r nrow(number1_median)` occupations in `r state_name` have the highest median annual pay compared to the other states in the country. Out of `r length(unique(state_wages$OCC_TITLE))`, that's `r round(length(number1_median)/length(unique(state_wages$OCC_TITLE))*100,0)` percent of all jobs with the highest pay categorized by the Bureau of Labor Statistics.

However, the value of a dollar differs state to state. Prices for some things can be cheaper or more expensive depending on where someone lives. For example, $`r dol_value1` in `r state_name` is actually $`r dol_value2` in `r dol_value2_state`, according to the Bureau of Economic Analysis.

So after adjusting for the actual value of a dollar in a given state, `r state_name` has about `r nrow(number1_median_adjusted)` jobs that have the highest median annual salary compared to other states.

Its lowest-paying jobs include `r paste0(min_job_median$OCC_TITLE[4], ", ", min_job_median$OCC_TITLE[3], ", and ", min_job_median$OCC_TITLE[2])`. After adjusting for cost of living, those occupational groups make $`r formatC(round(max(min_job_median$adjusted_A_MEAN[4])), format="d", big.mark=',')`,  $`r formatC(round(max(min_job_median$adjusted_A_MEAN[3])), format="d", big.mark=',')`, and  $`r formatC(round(max(min_job_median$adjusted_A_MEAN[2])), format="d", big.mark=',')`, respectively.

```{r time, include=F}
overtime <- read.csv("mega_wages.csv", stringsAsFactors=F)
# Employee growth over time

growth_no_na <-overtime %>%
  filter(!is.na(TOT_EMP)) %>%
  group_by(ST, OCC_CODE) %>%
  arrange(ST, OCC_CODE, YEAR) %>%
  filter(row_number()==1 | row_number()==n()) %>%
 mutate(TYPE=ifelse(row_number() %% 2 == 0, "year_last", "year_first")) %>%
 select(ST, STATE, OCC_CODE, TOT_EMP, YEAR, TYPE) 

growth_first <- growth_no_na %>%
  filter(TYPE=="year_first") 

growth_first$TYPE <- NULL
colnames(growth_first) <- c("ST", "STATE", "OCC_CODE", "FIRST_TOT_EMP", "FIRST_YEAR")


growth_last <- growth_no_na %>%
  filter(TYPE=="year_last") 
growth_last$TYPE <- NULL
colnames(growth_last) <- c("ST", "STATE", "OCC_CODE", "LAST_TOT_EMP", "LAST_YEAR")

growth_all <- left_join(growth_first, growth_last)
growth_all$percent_change_emp <- round((growth_all$LAST_TOT_EMP-growth_all$FIRST_TOT_EMP)/growth_all$FIRST_TOT_EMP*100,2)

growth <-overtime %>%
  group_by(ST, OCC_CODE) %>%
  select(ST, STATE, OCC_CODE, TOT_EMP, YEAR) %>%
  spread(YEAR, TOT_EMP)


titles <- read.csv("titles.csv", stringsAsFactors=F)
growth <- left_join(growth, titles)
growth$X <- NULL

growth_all <- left_join(growth_all, titles)
growth_all$X <- NULL
growth_all <- subset(growth_all, !is.na(LAST_TOT_EMP))


growth_all_state <- subset(growth_all, ST==state_abbr)
growth_all_state <- growth_all_state %>%
  filter(LAST_YEAR==2014 | LAST_YEAR==2015) %>%
  arrange(percent_change_emp)

growth_all_state_max <- tail(growth_all_state, 5)
growth_all_state_min <- head(growth_all_state, 5)

# Wage growth over time



wages_no_na <-overtime %>%
  filter(!is.na(adjusted_A_MEDIAN)) %>%
  group_by(ST, OCC_CODE) %>%
  arrange(ST, OCC_CODE, YEAR) %>%
  filter(row_number()==1 | row_number()==n()) %>%
 mutate(TYPE=ifelse(row_number() %% 2 == 0, "year_last", "year_first")) %>%
 select(ST, STATE, OCC_CODE, adjusted_A_MEDIAN, YEAR, TYPE) 

wages_first <- wages_no_na %>%
  filter(TYPE=="year_first") 

wages_first$TYPE <- NULL
colnames(wages_first) <- c("ST", "STATE", "OCC_CODE", "FIRST_adjusted_A_MEDIAN", "FIRST_YEAR")


wages_last <- wages_no_na %>%
  filter(TYPE=="year_last") 
wages_last$TYPE <- NULL
colnames(wages_last) <- c("ST", "STATE", "OCC_CODE", "LAST_adjusted_A_MEDIAN", "LAST_YEAR")

wages_all <- left_join(wages_first, wages_last)
wages_all$percent_change_wages <- round((wages_all$LAST_adjusted_A_MEDIAN-wages_all$FIRST_adjusted_A_MEDIAN)/wages_all$FIRST_adjusted_A_MEDIAN*100,2)

wages <-overtime %>%
  group_by(ST, OCC_CODE) %>%
  select(ST, STATE, OCC_CODE, adjusted_A_MEDIAN, YEAR) %>%
  spread(YEAR, adjusted_A_MEDIAN)


titles <- read.csv("titles.csv", stringsAsFactors=F)
wages <- left_join(wages, titles)
wages$X <- NULL

wages_all <- left_join(wages_all, titles)
wages_all$X <- NULL
wages_all <- subset(wages_all, !is.na(LAST_adjusted_A_MEDIAN))

wages_all_state <- subset(wages_all, ST==state_abbr)
wages_all_state <- wages_all_state %>%
  filter(LAST_YEAR==2014 | LAST_YEAR==2015) %>%
  arrange(percent_change_wages)

wages_all_state_max <- tail(wages_all_state, 5)
wages_all_state_min <- head(wages_all_state, 5)

```

`r wages_all_state_min$OCC_TITLE[1]` have seen the largest decline in wages. Between `r wages_all_state_min$FIRST_YEAR[1]` and `r wages_all_state_min$LAST_YEAR[1]`, the adjusted median annual salary dropped from $`r  formatC(round(min(wages_all_state_min$FIRST_adjusted_A_MEDIAN[1])), format="d", big.mark=',')` to $`r  formatC(round(min(wages_all_state_min$LAST_adjusted_A_MEDIAN[1])), format="d", big.mark=',')`. That's a change of about `r wages_all_state_min$percent_change_wages[1]` percent.

Meanwhile, `r str_to_lower(wages_all_state_max$OCC_TITLE[5])` saw the sharpest increase in wages in `r state_name`. Wages went up `r wages_all_state_max$percent_change_wages[5]` percent between `r wages_all_state_max$FIRST_YEAR[1]` and `r wages_all_state_max$LAST_YEAR[1]` from $`r  formatC(round(min(wages_all_state_max$FIRST_adjusted_A_MEDIAN[5])), format="d", big.mark=',')` to $`r  formatC(round(min(wages_all_state_max$LAST_adjusted_A_MEDIAN[5])), format="d", big.mark=',')`.

`r growth_all_state_min$OCC_TITLE[1]` lost the most workers in `r state_name`. In `r growth_all_state_min$FIRST_YEAR[1]`, there were `r  formatC(round(min(growth_all_state_min$FIRST_TOT_EMP[1])), format="d", big.mark=',')` employees. But by `r growth_all_state_min$LAST_YEAR[1]`, that figure declined `r growth_all_state_min$percent_change_emp[1]*-1` percent to `r  formatC(round(min(growth_all_state_min$LAST_TOT_EMP[1])), format="d", big.mark=',')`. 

The jobs that had gained the most employees was `r str_to_lower(growth_all_state_max$OCC_TITLE[5])`. There were `r formatC(round(min(growth_all_state_max$percent_change_emp[5])), format="d", big.mark=',')` percent more workers in `r growth_all_state_max$LAST_YEAR[1]` as compared to `r growth_all_state_max$FIRST_YEAR[1]`. Overall, the total number of workers grew from `r  formatC(round(min(growth_all_state_max$FIRST_TOT_EMP[5])), format="d", big.mark=',')` to `r  formatC(round(min(growth_all_state_max$LAST_TOT_EMP[5])), format="d", big.mark=',')` in `r state_name`.


```{r occu1, include=F}
occu_only_code_title <- subset(titles, OCC_TITLE==state_occ)
state_code <- occu_only_code_title$OCC_CODE

occu_only <- subset(wages2015, OCC_CODE==state_code)
occu_only_state <- subset(occu_only, ST==state_abbr)

occu_only <- occu_only %>%
  arrange(-adjusted_A_MEDIAN)

occu_only_low <- tail(occu_only, 5)
occu_only_high <- head(occu_only, 5)

the_wage <- occu_only_state$adjusted_A_MEAN

wage_sorted <- wages2015 %>%
  arrange(adjusted_A_MEDIAN)

wage_sorted_more <- subset(wage_sorted, adjusted_H_MEAN==the_wage & ST==state_abbr)
wage_sorted_more_row <- wage_sorted_more$row_num
wage_sorted_sorta <- filter(wage_sorted, ST!=state_abbr & adjusted_A_MEAN <= the_wage+10 & adjusted_A_MEAN >=the_wage-10)


r_n <- sample(1:nrow(wage_sorted_sorta), 1)


wages_all_state <- filter(wages_all, ST==state_abbr & OCC_CODE==state_code)
growth_all_state <- filter(growth_all, ST==state_abbr & OCC_CODE==state_code)

```

The adjusted median annual salary for `r str_to_lower(state_occ)` in `r state_name` is $`r formatC(round(min(occu_only_state$adjusted_A_MEDIAN)), format="d", big.mark=',')`, which is ranked `r occu_only_state$rank_adjusted_A_MEDIAN` in the country. `r ifelse(occu_only_state$rank_adjusted_A_MEDIAN==51, "", paste0(state_occ, " in ", occu_only_low$STATE[5], " make the lowest salary out of all the states at $", formatC(round(min(occu_only_low$adjusted_A_MEDIAN[5])), format="d", big.mark=','), "."))` `r ifelse(occu_only_state$rank_adjusted_A_MEDIAN==1, "", paste0(state_occ, " in ", occu_only_high$STATE[1], " make the highest salary at $", formatC(round(min(occu_only_high$adjusted_A_MEDIAN[1])), format="d", big.mark=','), "."))`

`r state_occ` in `r state_name` earn about the same as `r str_to_lower(wage_sorted_sorta$OCC_TITLE[r_n])` in `r wage_sorted_sorta$STATE[r_n]`.

The wages for `r str_to_lower(state_occ)` have `r ifelse(wages_all_state$FIRST_adjusted_A_MEDIAN > wages_all_state$LAST_adjusted_A_MEDIAN, "declined", "grown")` by `r ifelse(wages_all_state$percent_change_wages < 0, wages_all_state$percent_change_wages*-1, wages_all_state$percent_change_wages)` percent to $`r formatC(round(min(wages_all_state$LAST_adjusted_A_MEDIAN)), format="d", big.mark=',')` and the number of employees have `r ifelse(growth_all_state$FIRST_TOT_EMP > growth_all_state$LAST_TOT_EMP, "declined", "grown")` by `r ifelse(growth_all_state$percent_change_emp < 0, growth_all_state$percent_change_emp*-1, growth_all_state$percent_change_emp)` percent to `r formatC(round(min(growth_all_state$LAST_TOT_EMP)), format="d", big.mark=',')` in `r growth_all_state$LAST_YEAR`.